---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# measurementfree

<!-- badges: start -->
[![R build status](https://github.com/sociometricresearch/measurementfree/workflows/R-CMD-check/badge.svg)](https://github.com/sociometricresearch/measurementfree/actions)
[![Codecov test coverage](https://codecov.io/gh/sociometricresearch/measurementfree/branch/master/graph/badge.svg)](https://codecov.io/gh/sociometricresearch/measurementfree?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/measurementfree)](https://cran.r-project.org/package=measurementfree)
<!-- badges: end -->

The `measurementfree` package allows you to calculate several estimations of the quality of your survey questions and also adjust your estimations for measurement error.

Assuming you have a data frame with the `reliability`, `validity` and `quality` of your question, you can correct for the quality and common method variance easily.

Load the package as:

```{r}
# devtools::install_github("sociometricresearch/measurementfree")
library(measurementfree)
```

## A simple example

`measurementfree` introduces the concept of a measurement error design, such as the `survey` package has a survey design object. You can define this measurement error design with three objects: your model design, your measurement error data and the data of the analysis. For a simple case, let's use the data from the European Social Survey already loaded with the package. The variables `trstplt`, `trstprl` and `trstprt` were asked with the same type of the question. In other words, they share a common method (for example, a likert type scale question). Sharing a common method allows us to correct for their common method variance. We could define the measurement error design object as this:

```{r}
# Data
data(ess7es)
ess_subset <- ess7es[1:3]

# This is the model definition
model_definition <- "
  # Correct for measurement error
  ~~ trstplt + trstprl + trstprt

  # Correct for common method variance
  ~ trstplt + trstprl + trstprt
"

# The measurement error data
me_data <-
  data.frame(
    question = c("trstprl", "trstplt", "trstprt"),
    reliability = c(0.812, 0.852, 0.858),
    validity = c(0.959, 0.965, 0.956),
    quality = c(0.779, 0.822, 0.821)
  )

# Define your measurement error design
me_obj <- medesign(model_definition, ess_subset, me_data)
me_obj
```

This object describes your measurement error design. With this, we simply pass it to `me_cmv_cor` to adjust the correlation of `trstplt`, `trstprl` and `trstprt` for common method variance as well as their measurement error:

```{r }
me_cmv_cor(me_obj)
```

## Another simple example

The previous example corrected only variables `trstplt`, `trstprl` and `trstprt`. What if we have more groups of variables that share a common method? Just keep adding them to the model syntax! For example, `stfedu` and `stfhlth` also share a common method, let's add them to the model syntax and make sure we have data on `me_data` for them:

```{r }
# Data
data(ess7es)
ess_subset <- ess7es[1:5]

# This is the model definition
model_definition <- "
   # Correct for measurement error on all variables (.)
   ~~ .;

   # Correct for common method variance on these group of variables
   ~ trstplt + trstprl + trstprt;
   ~ stfedu + stfhlth
"

# The measurement error data
me_data <-
  data.frame(
    question = c("trstprl", "trstplt", "trstprt", "stfedu", "stfhlth"),
    reliability = c(0.812, 0.852, 0.858, 0.870, 0.871),
    validity = c(0.959, 0.965, 0.956, 0.915, 0.893),
    quality = c(0.779, 0.822, 0.821, 0.796, 0.779)
  )

# Define your measurement error design
me_obj <- medesign(model_definition, ess_subset, me_data)
me_obj
```

Once you have your measurement error design, we simply pass it to `me_cmv_cor` to adjust the correlation for common method variance as well as their quality:

```{r }
me_cmv_cor(me_obj)
```

Alternatively, you can use `me_cmv_cov` to adjust a covariance matrix for common method variance as well as quality.
